---
title: "Chapter 1"
output: html_notebook
---

Whoop whoop - a new R course: Much as I love coding in Python, RStudio is such a nice environment to explore - stronger than Jupyter for iterative data analysis.

```{r}
require(tidyverse)
```

```{r}
clvData1 <- read_csv('https://assets.datacamp.com/production/repositories/1861/datasets/4c612c0c990fccbd16c695043096aafbb66bd266/clvData1.csv')
```

```{r}
glimpse(clvData1)
```

```{r}
#install.packages("corrplot")
require(corrplot)
clvData1 %>%
  select(nOrders, nItems, daysSinceLastOrder, margin, returnRatio, shareOwnBrand,
         shareVoucher, shareSale, age, marginPerOrder, marginPerItem, itemsPerOrder, futureMargin) %>%
  cor() %>%
  corrplot()
```

```{r}
# data prep
salesData <- read_csv('https://assets.datacamp.com/production/course_6027/datasets/salesData.csv')
```


```{r}
# Structure of dataset
str(salesData, give.attr = FALSE)

# Visualization of correlations
salesData %>% select_if(is.numeric) %>%
  select(-id) %>%
  cor() %>%
  corrplot()

# Frequent stores
ggplot(salesData) +
    geom_boxplot(aes(x = mostFreqStore, y = salesThisMon))

# Preferred brand
ggplot(salesData) +
    geom_boxplot(aes(x = preferredBrand, y = salesThisMon))
```

```{r}
ggplot(clvData1, aes(margin, futureMargin)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE) +
  xlab("Margin year 1") +
  ylab("Margin year 2")
```

```{r}
# Model specification using lm
salesSimpleModel <- lm(salesThisMon ~ salesLast3Mon, 
                       data = salesData)

# Looking at model summary
summary(salesSimpleModel)

ggplot(salesData, aes(x=salesLast3Mon, y=salesThisMon)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE) +
  xlab("salesLast3Mon") +
  ylab("salesThisMon")
```

R squared can be worded as this explains 60% of the variation in the sales of this month. i.e. it's 40% distant/wrong!

```{r}
# what about a linear model on all the variables
multipleLM <- lm(futureMargin ~ ., data=clvData1)

summary(multipleLM)
```

But colinearity creates unstable co-efficients. e.g. nOrders and nItems are closely correlated.

```{r}
install.packages("rms")
library(rms)
vif(multipleLM)
```

Coefficients above 5 are colinear, and you should pick one and remove the other(s). Let's do that with nItems

```{r}
# data prep
cat(colnames(clvData1), sep=' + ')
```

```{r}
# what about a linear model on all the variables
multipleLM <- lm(futureMargin ~ nOrders + daysSinceLastOrder + margin +
                   returnRatio + shareOwnBrand + shareVoucher + shareSale + gender +
                   age + marginPerOrder + marginPerItem + itemsPerOrder, data=clvData1)

summary(multipleLM)
```

```{r}
# Estimating the full model
salesModel1 <- lm(salesThisMon ~ . - id, 
                 data = salesData)

# Checking variance inflation factors
vif(salesModel1)

# Estimating new model by removing information on brand
salesModel2 <- lm(salesThisMon ~ . - id - preferredBrand - nBrands, 
                 data = salesData)

# Checking variance inflation factors
vif(salesModel2)
```

```{r}
salesData2_4 <- read_csv('https://assets.datacamp.com/production/course_6027/datasets/salesDataMon2To4.csv')
```

```{r}
# getting an overview of new data
summary(salesData2_4)

# predicting sales
predSales5 <- predict(salesModel2, newdata = salesData2_4)

# calculating mean of future sales
mean(predSales5)
```



