---
title: "Chapter 4"
output: html_notebook
---

```{r}
require(tidyverse)
```

Whoop, whoop: Survival analysis: I need this for a current project.

Add a churn binary to ltv - tenure can remain the measure of time now, but churn=1 if end_date is less than now(). Then I need to use survival analysis to calculate recent tenure values.

```{r}
# data prep
data <- read_csv("https://assets.datacamp.com/production/course_6027/datasets/survivalDataExercise.csv")
dataNextOrder <- data[c("daysSinceFirstPurch", "boughtAgain")]
glimpse(dataNextOrder)
```


```{r}
# Look at the head of the data
head(dataNextOrder)

# Plot a histogram
ggplot(dataNextOrder) +
  geom_histogram(aes(x = daysSinceFirstPurch,
                     fill = factor(boughtAgain))) +
  facet_grid( ~ boughtAgain) + # Separate plots for boughtAgain = 1 vs. 0
  theme(legend.position = "none") # Don't show legend
```

### Survival Curve Analaysis = Kaplan Meier
*Survival function* - probability of 'death' versus time
*Hazard function* - the d/dt of a survival function. The peak is the highest occurence of 'deaths'

```{r}
# data prep
install.packages("survivql")
require(survival)
dataNextOrders <- data[c("daysSinceFirstPurch", "shoppingCartValue", "gender",
                         "voucher", "returned", "boughtAgain")]

# Create survival object
survObj <- Surv(dataNextOrder$daysSinceFirstPurch, dataNextOrder$boughtAgain)

# Look at structure
str(survObj)
```

```{r}
# data prep
dataNextOrder <- data[c("daysSinceFirstPurch", "shoppingCartValue", "gender",
                        "voucher", "returned", "boughtAgain")]

# Compute and print fit
fitKMSimple <- survfit(survObj ~ 1)
print(fitKMSimple)

# Plot fit
plot(fitKMSimple,
     conf.int = FALSE, xlab = "Time since first purchase",
     ylab = "Survival function", main = "Survival function")

# Compute fit with categorical covariate
fitKMCov <- survfit(survObj ~ voucher, data = dataNextOrder)

# Plot fit with covariate and add labels
plot(fitKMCov, lty = 2:3,
     xlab = "Time since first purchase", ylab = "Survival function", 
     main = "Survival function")
legend(90, .9, c("No", "Yes"), lty = 2:3)
```

### Cox Proportional Hazards

```{r}
library(rms)

# Determine distributions of predictor variables
dd <- datadist(dataNextOrder)
options(datadist = "dd")

# Compute Cox PH Model and print results
fitCPH <- cph(Surv(daysSinceFirstPurch, boughtAgain) ~ shoppingCartValue + voucher + returned + gender,
              data = dataNextOrder,
              x = TRUE, y = TRUE, surv = TRUE)
print(fitCPH)

# Interpret coefficients
print(exp(fitCPH$coefficients))

# Plot result summary
plot(summary(fitCPH), log = TRUE)
```

### Check model assumptions and make predictions
```{r}
# Check proportional hazard assumption and print result
testCPH <- cox.zph(fitCPH)
print(testCPH)

# Plot time-dependent beta
plot(testCPH, var = "gender=male")

# Load rms package
library(rms)

# Validate model
validate(fitCPH, method = "crossvalidation",
         B = 10, dxy = TRUE, pr = FALSE)
```

The R2 is the item to check here - and it is very poor: 0.03 - that's effectively useless at describing churn time with those variables. It needs more data - on our side intuition requires customer support tickets with a measure of unhappiness and perhaps external knowledge of searching for new services. i.e. we see it as a lead aggregator.

```{r}
# Create data with new customer
newCustomer <- data.frame(daysSinceFirstPurch = 21,
                  shoppingCartValue = 99.9,
                  gender = "female", 
                  voucher = 1,
                  returned = 0,
                  stringsAsFactors = FALSE)

# Make predictions
pred <- survfit(fitCPH, newdata = newCustomer)
print(pred)
plot(pred)

# Correct the customer's gender
newCustomer2 <- newCustomer
newCustomer2$gender <- "male"

# Redo prediction
pred2 <- survfit(fitCPH, newdata = newCustomer2)
print(pred2)
```






