---
title: "Machine Learning Toolbox"
output: html_notebook
---

# Chapter 2: Classification Models


```{r}
install.packages("caret", dependencies = TRUE)
```

```{r}
load("~/Documents/Code/R/datacamp/machine learning toolbox/Sonar.RData")
data(Sonar)
```

```{r}
# Get the Sonar data, and many more...

#install.packages("mlbench")
library(mlbench)
```

```{r}
data(Sonar)

# Shuffle row indices: rows
rows = sample(nrow(Sonar))

# Randomly order data: Sonar
Sonar <- Sonar[rows, ]

# Identify row to split on: split
split <- round(nrow(Sonar) * .60)

# Create train
train = Sonar[1:split, ]

# Create test
test = Sonar[(split + 1):nrow(Sonar), ]
```

```{r}
# Fit glm model: model
model <- glm(Class ~ ., family = "binomial", data = train)

# Predict on test: p
p <- predict(model, test, type = "response")
```

```{r}
library(caret)
# Calculate class probabilities: p_class
p_class <- ifelse(p > 0.50, "M", "R")

# Create confusion matrix
confusionMatrix(p_class, test$Class)
```

The model is terrible :) There are lots of false positive and negatives. It is actually worse; "No Information Rate : 0.5783" means we would get better outcomes if a monkey just said Mine all the time.

  * Sensitivy = how often we predict true correctly
  * Specificity = how often we predict negative correctly

```{r}
 (12 + 9) / sum(12,26,36,9)
```

```{r}
# Apply threshold of 0.9: p_class
p_class <- ifelse(p > 0.90, "M", "R")

# Create confusion matrix
confusionMatrix(p_class, test$Class)
```


```{r}
# Apply threshold of 0.10: p_class
p_class <- ifelse(p < 0.10, "R", "M")

# Create confusion matrix
confusionMatrix(p_class, test$Class)
```

### ROC curves

Iterate every possible classification threshold - which is then plotted.

```{r}
library(caTools)

# Predict on test: p
p <- predict(model, test, type = "response")

# Make ROC curve
colAUC(p, test$Class, plotROC = TRUE)
```

### AUC = Area Under Curve

AUC = area under the curve - 1 = perfect 0.5 = random guessing and 0.0 = perfectly wrong all the time! We just invert it to get perfectly right!

Generally want AUC 0.8 or higher, but 0.7 might be usuable

```{r}
# Create trainControl object: myControl
myControl <- trainControl(
  method = "cv",
  number = 10,
  summaryFunction = twoClassSummary,
  classProbs = TRUE, # IMPORTANT!
  verboseIter = TRUE
)
```

```{r}
train(Class ~ ., Sonar, method = "glm", trControl = myControl)

model
```









