---
title: "R Notebook"
output: html_notebook
---

# Exploratory Data Analysis in R

```{r}
library(dplyr)
votes <- readRDS("~/Documents/Code/R/datacamp/Exploratory Data Analysis/votes.rds")
descriptions <- readRDS("~/Documents/Code/R/datacamp/Exploratory Data Analysis/descriptions.rds")

votes
```

```{r}
#install.packages("countrycode")
#library(countrycode)

# Load the countrycode package
library(countrycode)

# Convert country code 100
countrycode(100, "cown", "country.name")

# Add a country column within the mutate: votes_processed
votes_processed <- votes %>%
  filter(vote <= 3) %>%
  mutate(year = session + 1945,
         country = countrycode(ccode, "cown", "country.name"))
```

```{r}
# You have the votes summarized by country
by_country <- votes_processed %>%
  group_by(country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# Print the by_country dataset
by_country

# Sort in ascending order of percent_yes
by_country %>%
  arrange(percent_yes)

# Now sort in descending order
by_country %>%
  arrange(desc(percent_yes))
```
```{r}
# Filter out countries with fewer than 100 votes
by_country %>%
  arrange(percent_yes) %>%
  filter(total >= 100)
```


```{r}
# Start with by_year_country dataset
by_year_country <- votes_processed %>%
  group_by(year, country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# Print by_year_country
by_year_country <- by_year_country %>% ungroup()
```

```{r}
library(ggplot2)

# Create a filtered version: UK_by_year
UK_by_year <- by_year_country %>%
  filter(country == "United Kingdom of Great Britain and Northern Ireland")

UK_by_year

# Line plot of percent_yes over time for UK only
ggplot(UK_by_year, aes(year, percent_yes)) +
  geom_line()
```

```{r}
# Vector of four countries to examine
countries <- c("United States of America", "United Kingdom of Great Britain and Northern Ireland",
               "France", "India")

# Filter by_year_country: filtered_4_countries
filtered_4_countries <- by_year_country %>%
  filter(country %in% countries)

# Line plot of % yes in four countries
ggplot(filtered_4_countries, aes(year, percent_yes, color = country)) +
  geom_line()
```

```{r}
# Line plot of % yes in four countries
ggplot(filtered_4_countries, aes(year, percent_yes)) +
  geom_line() +
  facet_grid(~ country)
```

```{r}
# Vector of six countries to examine
countries <- c("United States of America", "United Kingdom of Great Britain and Northern Ireland",
               "France", "Japan", "Brazil", "India")

# Filtered by_year_country: filtered_6_countries
filtered_6_countries <- by_year_country %>%
  filter(country %in% countries)

# Line plot of % yes over time faceted by country
ggplot(filtered_6_countries, aes(year, percent_yes, color = country)) +
  geom_line() +
  facet_wrap(~ country)
```


```{r}
# Line plot of % yes over time faceted by country
ggplot(filtered_6_countries, aes(year, percent_yes, color = country)) +
  geom_line() +
  facet_wrap(~ country, scales = "free_y")
```

```{r}
my_countries <- votes_processed %>%
  distinct(country) %>%
  arrange(country)
```


```{r}
countries <- c("United States of America", "United Kingdom of Great Britain and Northern Ireland",
               "France", "Japan", "Brazil", "India",
               "Ireland", "Romania", "Croatia")

# Filtered by_year_country: filtered_countries
filtered_countries <- by_year_country %>%
  filter(country %in% countries)

# Line plot of % yes over time faceted by country
ggplot(filtered_countries, aes(year, percent_yes, color = country)) +
  geom_line() +
  facet_wrap(~ country, scales = "free_y")
```

```{r}
# Percentage of yes votes from the US by year: US_by_year
US_by_year <- by_year_country %>%
  filter(country == "United States of America")

# Print the US_by_year data
US_by_year

# Perform a linear regression of percent_yes by year: US_fit
US_fit <- lm(percent_yes ~ year, data = US_by_year)

# Perform summary() on the US_fit object
summary(US_fit)
```

```{r}
library(broom)
tidy(US_fit)
```

```{r}
# Linear regression of percent_yes by year for US
US_by_year <- by_year_country %>%
  filter(country == "United States of America")
US_fit <- lm(percent_yes ~ year, US_by_year)

# Fit model for the United Kingdom
UK_by_year <- by_year_country %>%
  filter(country == "United Kingdom of Great Britain and Northern Ireland")
UK_fit <- lm(percent_yes ~ year, UK_by_year)

# Create US_tidied and UK_tidied
US_tidied <- tidy(US_fit)
UK_tidied <- tidy(UK_fit)

# Combine the two tidied models
bind_rows(US_tidied, UK_tidied)
```

```{r}
library(tidyr)

by_year_country %>%
  nest(-country)

by_year_country %>%
  nest(-country) %>%
  unnest(data)
```

```{r}
# All countries are nested besides country
nested <- by_year_country %>%
  nest(-country)

nested
nested$data[[7]]

# Unnest the data column to return it to its original form
nested %>%
  unnest(data)
```

```{r}
# Load tidyr and purrr
library(tidyr)
library(purrr)

# Perform a linear regression on each item in the data column
by_year_country %>%
  nest(-country) %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, .)))
```
```{r}
# Load the broom package
library(broom)

# Add another mutate that applies tidy() to each model
by_year_country %>%
  nest(-country) %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, .))) %>%
  mutate(tidied = map(model, tidy))
```

## Diagnosis

None of these transformations worked on my session, but they did in the course. This was because group is a setting which remains in a tibble for all subsequent transactions. When you print a tibble, any grouping is declared near the top, like below:

```{r}
> by_year_country
# A tibble: 4,744 x 4
# Groups:   year [34]
```

So in this case, nest obeys the group on year, which means year isn't available as a linear model factor. Calling ungroup into the same tibble removes the grouping, and nest works as expected by using the year in the nest, leaving the country as an index.

```{r}
# Print by_year_country
by_year_country <- by_year_country %>%
  ungroup()
```

```{r}
country_coefficients_all <- by_year_country %>%
  nest(-country) %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, .))) %>%
  mutate(tidied = map(model, tidy))

country_coefficients <- country_coefficients_all %>%
  select(country, tidied) %>%
  unnest(tidied)

# Filter by adjusted p-values
filtered_countries <- country_coefficients %>%
  filter(term == "year") %>%
  mutate(p.adjusted = p.adjust(p.value)) %>%
  filter(p.adjusted < .05)

# Sort for the countries increasing most quickly
filtered_countries %>%
  arrange(desc(statistic))

# Sort for the countries decreasing most quickly
filtered_countries %>%
  arrange(statistic)
```

## Joining and Tidying

```{r}
# Load dplyr package
library(dplyr)

# Print the votes_processed dataset
votes_processed

# Print the descriptions dataset
descriptions

# Join them together based on the "rcid" and "session" columns
# The error messages just states the table share nothing in common except their keys - which is fine!
votes_joined <- votes_processed %>%
  inner_join(descriptions, by = c("rcid", "session"))

votes_joined
```

```{r}
setequal(distinct(descriptions, rcid), distinct(votes_processed, rcid))
```

```{r}
votes_joined %>% filter(co > 0)
```

```{r}
# Load the ggplot2 package
library(ggplot2)

# Filter, then summarize by year: US_co_by_year
US_co_by_year <- votes_joined %>%
  filter(country == "United States of America" & co == 1) %>%
  group_by(year) %>%
  summarise(percent_yes = mean(vote == 1))

US_co_by_year

# Graph the % of "yes" votes over time
ggplot(US_co_by_year, aes(year, percent_yes)) +
  geom_line()
```

```{r}
# Load the tidyr package
library(tidyr)

# Gather the six me/nu/di/hr/co/ec columns
votes_joined %>%
  gather(topic, has_topic, me:ec)

# Perform gather again, then filter
# Gather the six me/nu/di/hr/co/ec columns
votes_gathered <- votes_joined %>%
  gather(topic, has_topic, me:ec) %>%
  filter(has_topic == 1)
```

```{r}
# Replace the two-letter codes in topic: votes_tidied
votes_tidied <- votes_gathered %>%
  mutate(topic = recode(topic,
                        me = "Palestinian conflict",
                        nu = "Nuclear weapons and nuclear material",
                        di = "Arms control and disarmament",
                        hr = "Human rights",
                        co = "Colonialism",
                        ec = "Economic development"))
```

```{r}
# Print votes_tidied
votes_tidied

# Summarize the percentage "yes" per country-year-topic
by_country_year_topic <- votes_tidied %>%
  group_by(country, year, topic) %>%
  summarize(total = n(), percent_yes = mean(vote == 1)) %>%
  ungroup()

# Print by_country_year_topic
by_country_year_topic
```

```{r}
# Load the ggplot2 package
library(ggplot2)

# Filter by_country_year_topic for just the US
US_by_country_year_topic <- by_country_year_topic %>%
  filter(country == "United States of America")

# Plot % yes over time for the US, faceting by topic
ggplot(US_by_country_year_topic, aes(year, percent_yes, color = topic)) +
  geom_line() +
  facet_wrap(~ topic)
```

```{r}
# Load purrr, tidyr, and broom
library(purrr)
library(tidyr)
library(broom)

# Print by_country_year_topic
by_country_year_topic

# Fit model on the by_country_year_topic dataset
country_topic_coefficients <- by_country_year_topic %>%
  nest(-country, -topic) %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, data = .)),
         tidied = map(model, tidy)) %>%
  unnest(tidied)

# Print country_topic_coefficients
country_topic_coefficients
```

```{r}
# Create country_topic_filtered
country_topic_filtered <- country_topic_coefficients %>%
  filter(term == "year") %>%
  mutate(p.adjusted = p.adjust(p.value)) %>%
  filter(p.adjusted < 0.05)

country_topic_filtered %>%
  filter(country %in% c("Afghanistan", "Malawi", "Vanuatu")) %>%
  select(country, topic, estimate) %>%
  arrange(estimate)
```

```{r}
# Create vanuatu_by_country_year_topic
vanuatu_by_country_year_topic <- by_country_year_topic %>%
  filter(country == "Vanuatu")

# Plot of percentage "yes" over time, faceted by topic
ggplot(vanuatu_by_country_year_topic, aes(year, percent_yes, col = topic)) +
  geom_line() +
  facet_wrap(~ topic)
```





















