---
title: "Chapter 2"
output: html_notebook
---

# Chapter 2

Epsilon defines how private we want to keep some information, and it creates a privacy budget. You reduce epsilon to increase the amount of data hiding required. There isn't a strict scale of epsilon - but the principle applies that large values leak data, and small hide data.

```{r}
load('data.RData')
```

## Global sensititivy 
Global sensitibity is the l1 distance of how the answer to a query changes if the most extreme case is present or absent from a dataset.

For example if Bill Gates attended a conference, this would not likely change the median salary of people that attended the conference, and therefore the global sensitivity would be small.

But if the mean salary were calculated, that would change a lot, and have a large global sensitivity: In this case, we would have to inject a lot of noise to protect the change that individual made to the answer.

```{r}
# Number of observations
n <- nrow(fertility)

# Global sensitivity of counts
gs.count <- 1

# Global sensitivity of proportions
gs.prop <- 1/n
```

```{r}
# Lower bound of Hours_Sitting
a <- 0

# Upper bound of Hours_Sitting
b <- 1

# Global sensitivity of mean for Hours_Sitting
gs.mean <- (b-a) / n

# Global sensitivity of proportions Hours_Sitting
gs.var <- (b-a)^2 / n
```

```{r}
library(dplyr)
library(smoothmest)

# How many participants had a Surgical_Intervention?
n_surgeries <- fertility %>%
    summarize_at(vars(Surgical_Intervention), sum) %>%
    pull()

n_surgeries

# Set the seed
set.seed(42)

# Apply the Laplace mechanism
eps <- 0.1
rdoublex(1, n_surgeries, 1/eps)
```

```{r}
# Proportion of Accident_Trauma
accident_prop <- fertility %>%
    summarise_at(vars(Accident_Trauma), mean) %>%
    pull()

accident_prop

# Set the seed
set.seed(42)

# Apply the Laplace mechanism
eps <- 0.1
rdoublex(1, accident_prop, gs.prop/eps)
```

```{r}
# Mean and Variance of Hours Sitting
hours_summary <- fertility %>%
    summarize_at(vars(Hours_Sitting), funs(mean, var))
hours_summary

# Setup
set.seed(42)
eps <- 0.1

# Laplace mechanism to mean
rdoublex(1, hours_summary$mean, gs.mean/eps)

# Laplace mechanism to variance
rdoublex(1, hours_summary$var, gs.var/eps)
```

