---
title: "Chapter 3"
output: html_notebook
---

## Sequential Composition
If you send the repated queries to the same dataset, the epsilon is divided amongst the repeating queries. e.g. mean and then max of dataset - epsilon is divided by two.

In reality - how do you predict that though?

```{r}
load('data.RData')
```

```{r}
# Load dplyr and smoothmest packages
library(dplyr)
library(smoothmest)

# Set Value of Epsilon
eps <- 0.1 / 2
```

```{r}
# Number of observations
n <- nrow(fertility)

# Lower bound of Age
a <- 0

# Upper bound of Age
b <- 1

# GS of counts for Diagnosis
gs.count <- 1

# GS of mean for Age
gs.mean <- (b-a)/n
```

```{r}
# Number of Participants with abnormal diagnosis
n_diagnosis <- fertility %>%
    summarise_at(vars(Diagnosis), sum)

n_diagnosis
	
# Mean of age
mean_age <- fertility %>%
    summarise_at(vars(Age), mean)

mean_age
```

```{r}
# Number of Participants with abnormal diagnosis
n_diagnosis <- fertility %>%
	summarize_at(vars(Diagnosis), sum) %>%
    pull()

# Mean of age
mean_age <- fertility %>%
	summarize_at(vars(Age), mean) %>%
    pull()

# Set seed
set.seed(42)

# Laplace mechanism to the count of abnormal diagnosis
rdoublex(1, n_diagnosis, 1/eps)

# Laplace mechanism to the mean of age
rdoublex(1, mean_age, gs.mean/eps)
```

## Parallel Composition
If multiple queries are sent to different parts of the dataset - then we do not need to divide epsilon. 

```{r}
# Set Value of Epsilon
eps <- 0.1

# Mean of Age per diagnosis level 
fertility %>%
    group_by(Diagnosis) %>%
    summarize_at(vars(Age), mean)
```

```{r}
# Set the seed
set.seed(42)

# Laplace mechanism to the mean age of participants with an abnormal diagnoisis
rdoublex(1, 0.71, gs.mean/eps)

# Laplace mechanism to the mean age of participants with a normal diagnoisis
rdoublex(1, 0.66, gs.mean/eps)
```

## Post-Processing
If you can answer a new query from differential data already provided - then you do not need to use the privacy budget. For instance, if you take the counts of smoking, which has three groups. No smoke, occasional, frequent. Then you can calculate the first two groups with differential privacy, and answer the final group by taking the first 2 from the total real count.

```{r}
# Set Value of Epsilon
eps <- 0.5 / (4-1)

# GS of Counts
gs.count <- 1

# Number of participants in each of the four seasons
fertility %>%
	count(Season)
```

```{r}
# Set the seed
set.seed(42)

# Laplace mechanism to the number of participants who were evaluated in the winter, spring, and summer
winter <- rdoublex(1, 28, gs.count / eps) %>%
  round()
	
spring <- rdoublex(1, 37, gs.count / eps) %>%
	round()
	
summer <- rdoublex(1, 4, gs.count / eps) %>%
	round()

# Post-process based on previous queries
fall <- nrow(fertility) - winter - spring - summer
```

## Impossible and Inconsistent Answers

```{r}
# Set Value of Epsilon
eps <- 0.01

# GS of counts
gs.count <- 1

# Number of Participants with Child_Disease
fertility %>%
	summarize_at(vars(Child_Disease), sum)
```

```{r}
# Apply the Laplace mechanism
set.seed(42)	
lap_childhood <- rdoublex(1, 87, gs.count / eps) %>%
	round()

# Total number of observations in fertility
max_value <- nrow(fertility)

# Bound the value such that the noisy answer does not exceed the total number of observations
ifelse(lap_childhood > max_value, max_value, lap_childhood)
```

```{r}
eps <- 0.005
gs.count <- 1

# Set the seed
set.seed(42)

# Apply the Laplace mechanism and avoid negative counts
fever1 <- rdoublex(1, 9, gs.count/eps) %>%
	max(0)
fever2 <- rdoublex(1, 63, gs.count/eps) %>%
	max(0)
fever3 <- rdoublex(1, 28, gs.count/eps) %>%
	max(0)
```

```{r}
fever <- c(fever1, fever2, fever3)

# Normalize noise 
fever_normalized <- (fever/sum(fever)) * (nrow(fertility))

# Round the values
round(fever_normalized)
```

