---
title: "R Notebook"
output: html_notebook
---

# Chapter 3

```{r}
library(data.table)
# The data.table package is pre-loaded

# Convert iris to a data.table
DT = data.table(iris)

# Species is "virginica"
DT[Species == "virginica"]

# Species is either "virginica" or "versicolor"
DT[Species %in% c("virginica", "versicolor")]
```

```{r}
# iris as a data.table
DT <- as.data.table(iris)

# Remove the "Sepal." prefix
old <- colnames(DT)
new <- gsub("Sepal.", "", colnames(DT))
setnames(DT, old, new)

# Remove the two columns starting with "Petal"
cols <- grep("Petal", colnames(DT))
DT <- DT[, !cols, with = FALSE]

colnames(DT)
```

With these exercises I tried a more complicated j = .(<blah) approach, which didn't work: The simple reference as a single match worked. This needs practice - data.table is very fast, but the syntax doesn't seem consistent: Practice will either show why that is the case by design, or learn the tricks through repetition.
```{r}
# Cleaned up iris data.table
DT

# Area is greater than 20 square centimeters
DT[Width * Length > 20]

# Add new boolean column
DT[, is_large := Width * Length > 25]
DT[, area := Width * Length]

# Now large observations with is_large
DT[is_large == TRUE]
```

I barely understood the last section. I needed to submit and use feedback to identify the right solution. I had forgotten about .N which is actually very useful: Getting the last row is VERY hard in a subset operation in base R.
```{r}
# The 'keyed' data.table DT
DT <- data.table(A = letters[c(2, 1, 2, 3, 1, 2, 3)], 
                 B = c(5, 4, 1, 9, 8, 8, 6), 
                 C = 6:12)
setkey(DT, A, B)

# Select the "b" group
DT["b"]

# "b" and "c" groups
DT[c("b", "c")]

# The first row of the "b" and "c" groups
DT[c("b", "c"), mult = "first"]

# First and last row of the "b" and "c" groups
DT[c("b", "c"), 
   .SD[c(1, .N)],
   mult = "all",
   by = .EACHI]

# Copy and extend code for instruction 4: add printout
DT[c("b", "c"), 
   { print(.SD); .SD[c(1, .N)]},
   mult = "all",
   by = .EACHI]
```

For these to make sense, you have to consider this is a form of data series imputing: That is the use case.

```{r}
# Keyed data.table DT
DT <- data.table(A = letters[c(2, 1, 2, 3, 1, 2, 3)], 
                 B = c(5, 4, 1, 9, 8, 8, 6), 
                 C = 6:12, 
                 key = "A,B")

# Get the key of DT
key(DT)

# Row where A == "b" and B == 6
DT[.("b", 6)]

# Return the prevailing row
DT[.("b", 6), roll = TRUE]

# Return the nearest row
DT[.("b", 6), roll = "nearest"]
```

```{r}
# Keyed data.table DT
DT <- data.table(A = letters[c(2, 1, 2, 3, 1, 2, 3)], 
                 B = c(5, 4, 1, 9, 8, 8, 6), 
                 C = 6:12, 
                 key = "A,B")

# Print the sequence (-2):10 for the "b" group
DT[.("b", (-2):10)]

# Add code: carry the prevailing values forwards
DT[.("b", (-2):10), roll = TRUE]

# Add code: carry the first observation backwards
DT[.("b", (-2):10), roll = TRUE, rollends = TRUE]

```


