---
title: "Exploratory Data Analysis"
output: html_notebook
---

levels(data$col) = uniq

on an entire set of cols
table(data) or table(data$col1, data$cal2)
equivalent to SQL SELECT count(col1), count(col1) FROM DATA GROUP BY col1, col2


This is a very pretty analysis of:

geom_density(fill = <facet>) 

```{r}
library(tidyverse)

glimpse(mtcars)

ggplot(mtcars, aes(x =  mpg, color = as.factor(cyl), fill = as.factor(cyl))) +
  geom_density(alpha = 0.5)
```

## Distance from Centre

var(x) = sample variance i.e. sum of square distance/ n - 1 (bessel correction for sample)

median and IQR are robust measures of distance from centre. They don't change much with outliers.

## Modality

unimodal
bimodal
multimodal
uniform

Skew = side of long tail

# Case Study

Examining an e-mail dataset, with measures for spam.

```{r}
# Load packages
library(ggplot2)
library(dplyr)
library(openintro)

# Compute summary statistics
email %>%
  group_by(spam) %>%
  summarize(
    median(num_char),
    IQR(num_char))

# Create plot
email %>%
  mutate(log_num_char = log(num_char)) %>%
  ggplot(aes(x = as.factor(spam), y = log_num_char)) +
  geom_boxplot()
  
email %>%
  mutate(log_num_char = log(num_char)) %>%
  ggplot(aes(x = log_num_char, fill = as.factor(spam))) +
  geom_density(alpha= 0.3)
```




