---
title: "R Core Plotting"
output: html_notebook
---

# Chapter 4

The ggplot course was a challenge - and at times I found the instructors soft voice deeply annoying. But the old boy in the core plotting course is great - clear, conscise, and delivers the connections at a pace I can connect in my brain whilst remaining engaged. He gave the best description of the QQ plot. Really like his pace, delivery and style. The plots are a bit old school - but its been great practice in helping stich all the new knowledge together.

```{r}
library(MASS)
#install.packages("car")
library(car)

# Set up a two-by-two plot array
par(mfrow = c(2,2))

# Plot the raw duration data
plot(geyser$duration, main = "Raw data")

# Plot the normalized histogram of the duration data
truehist(geyser$duration, main = "Histogram")

# Plot the density of the duration data
plot(density(geyser$duration), main = "Density")

# Construct the normal QQ-plot of the duration data
qqPlot(geyser$duration, main = "QQ-plot")
```

```{r}
# Load the insuranceData package
#install.packages("insuranceData")
library(insuranceData)

# Use the data() function to load the dataCar data frame
data(dataCar)

# Load the tabplot package
#install.packages("tabplot")
suppressPackageStartupMessages(library(tabplot))

str(dataCar)

# Generate the default tableplot() display
tableplot(dataCar)
```





