---
title: "R Notebook"
output: html_notebook
---

# Chapter 2

```{r}
# unemployment is in the workspace
summary(unemployment)

# unemployment_model is in the workspace
summary(unemployment_model)

# Make predictions from the model
unemployment$predictions <- predict(unemployment_model)

# Fill in the blanks to plot predictions (on x-axis) versus the female_unemployment rates
ggplot(unemployment, aes(x = predictions, y = female_unemployment)) + 
  geom_point() + 
  geom_abline()

# Calculate residuals
unemployment$residuals <- residuals(unemployment_model)

# Fill in the blanks to plot predictions (on x-axis) versus the residuals
ggplot(unemployment, aes(x = predictions, y = residuals)) + 
  geom_pointrange(aes(ymin = 0, ymax = residuals)) + 
  geom_hline(yintercept = 0, linetype = 3) + 
  ggtitle("residuals vs. linear model prediction")
```

```{r}
# unemployment is in the workspace (with predictions)
summary(unemployment)

# unemployment_model is in the workspace
summary(unemployment_model)

# Load the package WVPlots
#install.packages("WVPlots")
library(WVPlots)

# Plot the Gain Curve
GainCurvePlot(unemployment, "predictions"
              , "female_unemployment",  "Unemployment model")
```

```{r}
# unemployment is in the workspace
summary(unemployment)

# For convenience put the residuals in the variable res
res <- unemployment$residuals

# Calculate RMSE, assign it to the variable rmse and print it
(rmse <- sqrt(mean(res^2)))

# Calculate the standard deviation of female_unemployment and print it
(sd_unemployment <- sd(unemployment$female_unemployment))
```

```{r}
# unemployment is in your workspace
summary(unemployment)

# unemployment_model is in the workspace
summary(unemployment_model)

# Calculate mean female_unemployment: fe_mean. Print it
(fe_mean <- mean(unemployment$female_unemployment))

# Calculate total sum of squares: tss. Print it
(tss <- sum((fe_mean - unemployment$female_unemployment)^2))

# Calculate residual sum of squares: rss. Print it
(rss <- sum(unemployment$residuals^2))

# Calculate R-squared: rsq. Print it. Is it a good fit?
(rsq <- 1 - rss/tss)

# Get R-squared from glance. Print it
(rsq_glance <- glance(unemployment_model)$r.squared)

# So if our model has a great fit, the resisuals are small
# which makes R^2 close to 1.
# but it if the residuals are large, and in the worst case
# identical to our variance - then R^2 is close to zero
```

The linear correlation of two variables, xx and yy, measures the strength of the linear relationship between them. When xx and yy are respectively:

   * the outcomes of a regression model that minimizes squared-error (like linear regression) and
   * the true outcomes of the training data,

then the square of the correlation is the same as R2R2. You will verify that in this exercise.

```{r}
# unemployment is in your workspace
summary(unemployment)

# unemployment_model is in the workspace
summary(unemployment_model)

# Get the correlation between the prediction and true outcome: rho and print it
(rho <- cor(unemployment$female_unemployment, unemployment$predictions))

# Square rho: rho2 and print it
(rho2 <- rho^2)

# Get R-squared from glance and print it
(rsq_glance <- glance(unemployment_model)$r.squared)
```

